Describing vocalizations in young children: A big data approach through citizen science annotation

### 2021-02-18 202010_jslhr/paper.Rmd
ready for final proofing

#fix typo
chunks$segmentId_DB=gsub("(DB.)(.)$","\\1",chunks$segmentId_DB)

@@

label_options=c("Canonical"   ,     "Crying"  ,        "Junk"  ,    "Laughing", "Non-Canonical")

@@

Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations. @cychosz2019canonical found that the Canonical Proportion extracted from daylong recordings was significantly correlated with age in a multicultural and multilingual sample of children up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations. That study built on a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this prior work focuses on babbling and thus typically on infants under one year of age. In these younger infants, a critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech [@oller1998late]. Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech. When compared to more primitive sounds such as squeals, or isolated vowels, canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa). Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence;@cychosz2019canonical]. In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder [@mcdaniel2020predicting].

@@

Citizen science may be particularly helpful when analyzing infants' data from wearables, which remain challenging to annotate as mentioned above. There is one previous study that attempted this approach. @cychosz2019canonical drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî. For the English and Tsimane' corpora, vocalizations were automatically identified using LENA&reg;, whereas the other two were extracted through manual segmentation. Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice). They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgment of canonical or non-canonical. They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables. The authors also studied how their Canonical Proportion related to age within each corpus that had a sufficient number of children varying in age. They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "Junk" category was very high. Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA&reg; than through manual segmentation. In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.

@@

Our work seeks to broaden the already promising results that have emerged from previous work summarized above. We hope that this methodology will open the road to larger scale analyses of children's vocalizations as captured by wearables. In addition, we went beyond @cychosz2019canonical's study in two important ways. First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion. In contrast, @cychosz2019canonical actively recruited people to provide annotations, stating that "[a]nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators". A process that draws from research assistants relies on expert resources, whose time is expensive. Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework. Those authors also had to rely on manual extraction of the resulting data from the iHearUplay team, which added to the administrative load. Second, a core goal of the present work was to determine how citizen scientists' annotations fare compared to the current gold standard, laboratory annotations. @cychosz2019canonical did not have a gold standard for their data.

@@

It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels. We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA&reg; [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range. Some previous work had shown that vocal development proceeded differently when compared to a normative sample [@hamrick2019capturing;@hamrick2019measurement]. Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables [@rankine2017language;@belardi2017retrospective; see also @roche2018early for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics]. Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods [see @rankine2017language for discussion], since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording). Additionally, the ideal method will be sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children. Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.

@@

Specifically, the segments extracted were automatically cut into clips of exactly 500 ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform. We will call this level *chunk*, to convey that they are part of a bigger unit (the LENA&reg;-identified "segment"). Note that segments and chunks are collectively referred to as *clips* because they are parts of the longer audio recording. To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500-ms chunks and their corresponding segment. Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].

Citizen scientists could access our project, "Maturity of Baby Sounds", by logging in, or as anonymous participants. Before starting the annotation, participants were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1). In this tutorial, we included one audio example for each category in order to make the classification task as smooth as possible. Further clarifications on what constitutes canonical and non-canonical sequences, as well as five audio examples for each of the categories, were available through the Field Guide, which users could access by clicking on the right side of the screen. The Field Guide could be consulted at any point of the classification without interrupting the task.

@@

<!-- lh said: I'm also wondering what the breakdown of lab annotations was for the Zooniverse segments without majority agreement -->

@@

This is essentially the same thought process Lab annotators followed when jduging segments: a LENA&reg; segment may contain 1 or more vocalizations, which may be of the same type or not. Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels.

@@

```{r clean up, echo=F}
# remove non-majority labels from lab
data_all[data_all$Num_Agreement>2,]->data_all

# remove non-majority labels from zooniverse
data_all[data_all$Zoon_classif %in% label_options,]->data_all

table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100)
table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100)
```

@@

@@

}
```

@@

Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups. There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r printr(myr["l_lp","AS"])`, CI [`r printr(mylb["l_lp","AS"])`,`r printr(myhb["l_lp","AS"])`], *p* = `r printp(myp["l_lp","AS"])`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r printr(myr["l_lp","LR"])`, CI [`r printr(mylb["l_lp","LR"])`,`r printr(myhb["l_lp","LR"])`], *p* = `r printp(myp["l_lp","LR"])`]. The Canonical Proportion exhibited non-significant developmental decreases among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r printr(myr["l_cp","AS"])`, CI [`r printr(mylb["l_cp","AS"])`,`r printr(myhb["l_cp","AS"])`], *p* = `r printp(myp["l_cp","AS"])`]; and marginal developmental increases among low-risk control r(`r myn["l_cp","LR"]`) = `r printr(myr["l_cp","LR"])`, CI [`r printr(mylb["l_cp","LR"])`,`r printr(myhb["l_cp","LR"])`], *p* = `r printp(myp["l_cp","LR"])`].

Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r printr(myr["z_lp","AS"])`, CI [`r printr(mylb["z_lp","AS"])`,`r printr(myhb["z_lp","AS"])`], *p* = `r printp(myp["z_lp","AS"])`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r printr(myr["z_lp","LR"])`, CI [`r printr(mylb["z_lp","LR"])`,`r printr(myhb["z_lp","LR"])`], *p* = `r printp(myp["z_lp","LR"])`]. Similarly, there were non-significant developmental decreases in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r printr(myr["z_cp","AS"])`, CI [`r printr(mylb["z_cp","AS"])`,`r printr(myhb["z_cp","AS"])`], *p* = `r printp(myp["z_cp","AS"])`]; and marginal developmental increases among low-risk control children, r(`r myn["z_cp","LR"]`) = `r printr(myr["z_cp","LR"])`, CI [`r printr(mylb["z_cp","LR"])`,`r printr(myhb["z_cp","LR"])`], *p* = `r printp(myp["z_cp","LR"])`].

@@

Next, we discuss the correspondence between citizen science classifications and the laboratory gold standard, at the level of individual segments. Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level classifications for each class while the off-diagonal elements show non-matching classifications.

@@

mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
#mycf
```

@@

These visualizations suggest that performance is moderate to good, which was confirmed via statistical analyses. The overall (weighted) accuracy is `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r printr(mycf$overall["Kappa"])`, and the Gwet's AC1 coefficient is `r printr(mygac$est)`, CI = [`r printr(mygac$lb)`,`r printr(mygac$ub)`].

@@

mycortab=t(myr)
rownames(mycortab)<-paste("All seg",rownames(mycortab))
```

Although the classification at the segment level is only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development. We illustrate this association in Figure 5. Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.  When we split by participant group, correlations remain high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`]; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`].

@@

Similarly, a strong positive correlation is found in the Canonical Proportion r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`]. When we split by participant group, correlations remain high although we do note they are somewhat smaller for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.

@@

In this section, we report on explorations of under what conditions Zooniverse labels more closely aligned with laboratory labels. In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2019canonical]. A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.

@@

      #get cor properties
  thiscor=cor.test(proportions[proportions$median_Junk!="lower_Junk","l_cp"],proportions[proportions$median_Junk!="lower_Junk","z_cp"])
    myr[2,"higher_Junk"]=round(thiscor$estimate,3)
    mylb[2,"higher_Junk"]=round(thiscor$conf.int[1],3)
    myhb[2,"higher_Junk"]=round(thiscor$conf.int[2],3)
    myn[2,"higher_Junk"]=round(thiscor$parameter,3)
    myp[2,"higher_Junk"]=round(thiscor$p.value,3)
```

We investigated this hypothesis by calculating the proportion of their data labeled as "Junk" for each individual child.  There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"],3)`) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"],3)`): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, *p* = `r printp(JunkXdiag$p.value)`. We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk". Results were similar across these two post-hoc subgroups (see Figure 6). For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r printr(myr["lp","lower_Junk"])`, CI [`r printr(mylb["lp","lower_Junk"])`,`r printr(myhb["lp","lower_Junk"])`], *p* = `r printp(myp["lp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r printr(myr["lp","higher_Junk"])`, CI [`r printr(mylb["lp","higher_Junk"])`,`r printr(myhb["lp","higher_Junk"])`], *p* `r printp(myp["lp","higher_Junk"])`. For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r printr(myr["cp","lower_Junk"])`, CI [`r printr(mylb["cp","lower_Junk"])`,`r printr(myhb["cp","lower_Junk"])`], *p* `r printp(myp["cp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r printr(myr["cp","higher_Junk"])`, CI [`r printr(mylb["cp","higher_Junk"])`,`r printr(myhb["cp","higher_Junk"])`], *p* `r printp(myp["cp","higher_Junk"])`. Thus, it does not seem that a higher proportion of "Junk" labels is an index of low quality data.

@@

Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level labels) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse participants) to segments (the original LENA&reg; segments presented to laboratory participants), because one can calculate the proportions directly from the chunk-level judgment. As a result, there was one stage in our pre-processing that may not have been necessary, whereby we collapsed labels across chunks associated to the same segment. We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations. To facilitate comparison, correlations across laboratory- and Zooniverse-derived proportions are provided in Table 2.

@@

>>>>

@@

label_options=c("Canonical" , "Non-Canonical"  ,     "Crying"   ,    "Laughing",        "Junk" )

@@

Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations. @cychosz2021canonical found that the Canonical Proportion extracted from daylong recordings was significantly correlated with age in a multicultural and multilingual sample of children up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations. That study built on a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this prior work focuses on babbling and thus typically on infants under one year of age. In these younger infants, a critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech [@oller1998late]. Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech. When compared to more primitive sounds such as squeals, or isolated vowels, canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa). Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development [@oller2000emergence;@cychosz2021canonical]. In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder [@mcdaniel2020predicting].

@@

Citizen science may be particularly helpful when analyzing infants' data from wearables, which remain challenging to annotate as mentioned above. There is one previous study that attempted this approach. @cychosz2021canonical drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî. For the English and Tsimane' corpora, vocalizations were automatically identified using LENA&reg;, whereas the other two were extracted through manual segmentation. Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice). They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgment of canonical or non-canonical. They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables. The authors also studied how their Canonical Proportion related to age within each corpus that had a sufficient number of children varying in age. They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the "Junk" category was very high. Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA&reg; than through manual segmentation. In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.

@@

Our work seeks to broaden the already promising results that have emerged from previous work summarized above. We hope that this methodology will open the road to larger scale analyses of children's vocalizations as captured by wearables. In addition, we went beyond @cychosz2021canonical's study in two important ways. First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion. In contrast, @cychosz2021canonical actively recruited people to provide annotations, stating that "[a]nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators". A process that draws from research assistants relies on expert resources, whose time is expensive. Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework. Second, a core goal of the present work was to determine how citizen scientists' annotations fare compared to the current gold standard, laboratory annotations. @cychosz2021canonical did not have a gold standard for their data.

@@

It was important for generalizability purposes to examine data from children with variable biological age, mental age, and vocal maturity levels. We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA&reg; [see meta-analyses in @wang2020meta;@cristia2020accuracy]; as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range. Some previous work had shown that vocal development proceeded differently among children with an Angelman syndrome diagnosis when compared to a normative sample [@hamrick2019capturing;@hamrick2019measurement]. Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables [@rankine2017language;@belardi2017retrospective; see also @roche2018early for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics]. Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods [see @rankine2017language for discussion], since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording). Additionally, the ideal method is sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children. Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.

@@

Specifically, the segments extracted were automatically cut into clips of exactly 500 ms, extracting neighboring silence when necessary, before being uploaded on the Zooniverse platform. We call this level *chunk*, to convey that they are part of a bigger unit (the LENA&reg;-identified "segment"). Note that segments and chunks are collectively referred to as *clips* because they are parts of the longer audio recording. To allow the recovery of the original segments at later stages of the analysis, the scripts created a metadata file with the mapping between 500-ms chunks and their corresponding segment. Chunks were then uploaded on Zooniverse using Panoptes, their open-source, command-line based API for data handling [@bowyer2015panoptes].

Citizen scientists could access our project, "Maturity of Baby Sounds", by logging in, or as anonymous annotators. Before starting the annotation, citizen scientists were given a quick tutorial which walked them through the steps of the annotation workflow (i.e., how to play a sound, how to make a selection; see Figure 1). In this tutorial, we included one audio example for each category in order to make the classification task as smooth as possible. Further clarifications on what constitutes canonical and non-canonical sequences, as well as five audio examples for each of the categories, were available through the Field Guide, which users could access by clicking on the right side of the screen. The Field Guide could be consulted at any point of the classification without interrupting the task.

@@

```{r clean up, echo=F}
#dim(data_all) #11980    16
# remove non-majority labels from lab
data_all[data_all$Num_Agreement>0,]->data_all
#dim(data_all) # 11647    16

# remove non-majority labels from zooniverse
data_all[data_all$Zoon_classif %in% label_options,]->data_all
#dim(data_all) # 11268    16

table_zoon=round(table(data_all$Zoon)/sum(table(data_all$Zoon))*100,2)
table_lab=round(table(data_all$lab)/sum(table(data_all$lab))*100,2)
```

@@

This is essentially the same thought process Lab annotators followed when jduging segments: a LENA&reg; segment may contain 1 or more vocalizations, which may be of the same type or not. Figure 2 conveys the relationship between chunks and segments, and how judgments are combined into labels. After removing segments for which there was no majority label in the lab and/or Zooniverse coding, we were left with data for `r dim(data_all)[1]` segments.

@@

@@

  #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis=="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis=="AngelmanSyndrome"])
    myr[thisvar,"AS"]=printr(thiscor$estimate)
    mylb[thisvar,"AS"]=round(thiscor$conf.int[1],3)
    myhb[thisvar,"AS"]=round(thiscor$conf.int[2],3)
    myn[thisvar,"AS"]=round(thiscor$parameter,3)
    myp[thisvar,"AS"]=round(thiscor$p.value,3)

@@

      #get cor properties
  thiscor=cor.test(proportions[proportions$Diagnosis!="AngelmanSyndrome",thisvar],proportions$Age[proportions$Diagnosis!="AngelmanSyndrome"])
    myr[thisvar,"LR"]=printr(thiscor$estimate)
    mylb[thisvar,"LR"]=round(thiscor$conf.int[1],3)
    myhb[thisvar,"LR"]=round(thiscor$conf.int[2],3)
    myn[thisvar,"LR"]=round(thiscor$parameter,3)
    myp[thisvar,"LR"]=round(thiscor$p.value,3)

}

#matrix that will hold all r results
corage<-t(myr)[,c("l_lp","z_lp","l_cp","z_cp")]

@@

Descriptive analyses on the laboratory annotations showed that correlations between the Linguistic Proportion and age differed across the groups. There was a near-zero relationship among the children diagnosed with Angelman syndrome r(`r myn["l_lp","AS"]`) = `r (myr["l_lp","AS"])`, CI [`r printr(mylb["l_lp","AS"])`,`r printr(myhb["l_lp","AS"])`], *p* = `r printp(myp["l_lp","AS"])`]; and a significant association among low-risk control children r(`r myn["l_lp","LR"]`) = `r (myr["l_lp","LR"])`, CI [`r printr(mylb["l_lp","LR"])`,`r printr(myhb["l_lp","LR"])`], *p* = `r printp(myp["l_lp","LR"])`]. The Canonical Proportion exhibited a non-significant developmental decrease among children diagnosed with Angelman syndrome r(`r myn["l_cp","AS"]`) = `r (myr["l_cp","AS"])`, CI [`r printr(mylb["l_cp","AS"])`,`r printr(myhb["l_cp","AS"])`], *p* = `r printp(myp["l_cp","AS"])`]; and a marginal developmental increase among low-risk control r(`r myn["l_cp","LR"]`) = `r (myr["l_cp","LR"])`, CI [`r printr(mylb["l_cp","LR"])`,`r printr(myhb["l_cp","LR"])`], *p* = `r printp(myp["l_cp","LR"])`].

Using the Zooniverse annotations, we found that the association with age was very weak for children diagnosed with Angelman syndrome r(`r myn["z_lp","AS"]`) = `r (myr["z_lp","AS"])`, CI [`r printr(mylb["z_lp","AS"])`,`r printr(myhb["z_lp","AS"])`], *p* = `r printp(myp["z_lp","AS"])`]; whereas low-risk control children showed a significant increase with age r(`r myn["z_lp","LR"]`) = `r (myr["z_lp","LR"])`, CI [`r printr(mylb["z_lp","LR"])`,`r printr(myhb["z_lp","LR"])`], *p* = `r printp(myp["z_lp","LR"])`]. Similarly, there was a non-significant developmental decrease in the Canonical Proportion among children with Angelman syndrome r(`r myn["z_cp","AS"]`) = `r (myr["z_cp","AS"])`, CI [`r printr(mylb["z_cp","AS"])`,`r printr(myhb["z_cp","AS"])`], *p* = `r printp(myp["z_cp","AS"])`]; and a marginal developmental increase among low-risk control children, r(`r myn["z_cp","LR"]`) = `r (myr["z_cp","LR"])`, CI [`r printr(mylb["z_cp","LR"])`,`r printr(myhb["z_cp","LR"])`], *p* = `r printp(myp["z_cp","LR"])`].

@@

Next, we discuss the correspondence between citizen science labels and the laboratory gold standard, at the level of individual segments. Results were visualized with a confusion matrix showing precision and recall (Figure 4): the diagonal elements show the number of correct segment-level labels for each class while the off-diagonal elements show non-matching labels.

@@

mycf=confusionMatrix(data_all$lab, data_all$Zoon_classif, dnn = c("Lab","Zooniverse"))
conf_tab=mycf$table
# this package uses sensitivity & specificity
#Sensitivity=recall
#Specificity=precision
#mycf

#compose precision
colsums=colSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/colsums[i]
#colSums(my_conf_tab) #check
prop_cat=data.frame(my_conf_tab*100) #generates precision because columns
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"pr"
data.frame(conf_tab)->stall_p
stall_p$id=paste(stall_p$Lab,stall_p$Zooniverse)
stall_p=merge(stall_p,prop_cat[c("id","pr")])

# repeat for recall
rowsums=rowSums(conf_tab)
my_conf_tab=conf_tab
for(i in 1:dim(conf_tab)[2]) my_conf_tab[,i]=my_conf_tab[,i]/rowsums[i]
#rowSums(my_conf_tab)
prop_cat=data.frame(conf_tab/rowSums(conf_tab)*100)  #generates recall because rows
prop_cat$id=paste(prop_cat$Lab,prop_cat$Zooniverse)
colnames(prop_cat)[3]<-"rec"
data.frame(conf_tab)->stall_r
stall_r$id=paste(stall_r$Lab,stall_r$Zooniverse)
stall_r=merge(stall_r,prop_cat[c("id","rec")])
```

@@

These visualizations suggest that performance was moderate to good, which was confirmed via statistical analyses. The overall (weighted) accuracy was `r round(mycf$overall["Accuracy"]*100)`%, CI = [`r round(mycf$overall["AccuracyLower"]*100)`,`r round(mycf$overall["AccuracyUpper"]*100)`],  kappa is `r printr(mycf$overall["Kappa"])`, and the Gwet's AC1 coefficient is `r printr(mygac$est)`, CI = [`r printr(mygac$lb)`,`r printr(mygac$ub)`].

@@

mycortab=t(myr)
rownames(mycortab)<-paste("All seg",rownames(mycortab))

Although the classification at the segment level was only moderately accurate, what we are ultimately interested in is whether citizen scientists’ classifications are able to provide a reliable snapshot of children's individual development. We illustrate this association in Figure 5. Looking at all 20 children together, we found a strong positive correlation r(`r myn["lp","all"]`) = `r printr(myr["lp","all"])`, CI [`r printr(mylb["lp","all"])`,`r printr(myhb["lp","all"])`], *p* `r printp(myp["lp","all"])`] between Linguistic Proportion by child from the Zooniverse and the lab annotators' data.  When we split by participant group, correlations remained high: for Angelman syndrome r(`r myn["lp","AS"]`) = `r printr(myr["lp","AS"])`, CI [`r printr(mylb["lp","AS"])`,`r printr(myhb["lp","AS"])`], *p* = `r printp(myp["lp","AS"])`]; low-risk control r(`r myn["lp","LR"]`) = `r printr(myr["lp","LR"])`, CI [`r printr(mylb["lp","LR"])`,`r printr(myhb["lp","LR"])`], *p* = `r printp(myp["lp","LR"])`].

@@

Similarly, a strong positive correlation was found in the Canonical Proportion r(`r myn["cp","all"]`) = `r printr(myr["cp","all"])`, CI [`r printr(mylb["cp","all"])`,`r printr(myhb["cp","all"])`], *p* `r printp(myp["cp","all"])`]. When we split by participant group, correlations remained high although we do note they were somewhat lower for the children with Angelman syndrome: r(`r myn["cp","AS"]`) = `r printr(myr["cp","AS"])`, CI [`r printr(mylb["cp","AS"])`,`r printr(myhb["cp","AS"])`], *p* = `r myp["cp","AS"]`]; than the low-risk control children r(`r myn["cp","LR"]`) = `r printr(myr["cp","LR"])`, CI [`r printr(mylb["cp","LR"])`,`r printr(myhb["cp","LR"])`], *p* `r printp(myp["cp","LR"])`.

@@

In this section, we report on exploratory analyses, aimed at establishing conditions under which Zooniverse labels aligned with laboratory labels more or less closely. In previous work using a similar method, for instance, data from all three children from one dataset were often labeled as "Junk" (i.e., not a child's vocalization), and the data points from this corpus stood out when the authors attempted to integrate results with other corpora [@cychosz2021canonical]. A high proportion of "Junk" may indicate that automated segmentation was errorful for those children, and may be a sign that the rest of the data could be compromised as well.

@@

We investigated this hypothesis by calculating the proportion of data labeled as "Junk" for each individual child.  There was no significant difference in the proportions of their data labeled as "Junk" for children with Angelman syndrome (M = `r round(JunkXdiag$estimate["mean in group AngelmanSyndrome"]*100,2)`%, SD = `r round(sd(proportions[proportions$Diagnosis=="AngelmanSyndrome","z_Junk"])*100,2)`%) compared to the low-risk group (M = `r round(JunkXdiag$estimate["mean in group Low-RiskControl"]*100,2)`%, SD = `r round(sd(proportions[proportions$Diagnosis!="AngelmanSyndrome","z_Junk"])*100,2)`%): Welch's t(`r JunkXdiag$parameter`) = `r round(JunkXdiag$statistic,3)`, *p* = `r printp(JunkXdiag$p.value)`. We therefore collapsed across groups for this exploratory analysis, and split the 20 children using a median split on the proportion of their data labeled as "Junk". Results were similar across these two post-hoc subgroups (see Figure 6). For Linguistic Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["lp","lower_Junk"]`) = `r printr(myr["lp","lower_Junk"])`, CI [`r printr(mylb["lp","lower_Junk"])`,`r printr(myhb["lp","lower_Junk"])`], *p* = `r printp(myp["lp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["lp","higher_Junk"]`) = `r printr(myr["lp","higher_Junk"])`, CI [`r printr(mylb["lp","higher_Junk"])`,`r printr(myhb["lp","higher_Junk"])`], *p* `r printp(myp["lp","higher_Junk"])`. For Canonical Proportion, the correlation across lab and Zooniverse data for the lower Junk group was r(`r myn["cp","lower_Junk"]`) = `r printr(myr["cp","lower_Junk"])`, CI [`r printr(mylb["cp","lower_Junk"])`,`r printr(myhb["cp","lower_Junk"])`], *p* `r printp(myp["cp","lower_Junk"])`]; and for the higher Junk group it was r(`r myn["cp","higher_Junk"]`) = `r printr(myr["cp","higher_Junk"])`, CI [`r printr(mylb["cp","higher_Junk"])`,`r printr(myhb["cp","higher_Junk"])`], *p* `r printp(myp["cp","higher_Junk"])`. Thus, it does not seem that a higher proportion of "Junk" labels is an index of low quality data.

@@

Above, we concluded that derived metrics integrating information across audio clips (Linguistic and Canonical Proportion, which can be derived from segment- or chunk-level labels) seem more promising than segment-level data (where individual segments are classified into Crying, Laughing, etc.) Notice that our derived metrics do not require matching of chunks (500 ms presented to Zooniverse annotators) to segments (the original LENA&reg; segments presented to laboratory annotators), because one can calculate the proportions directly from the chunk-level judgment. As a result, there was one stage in our post-processing that may not have been necessary, whereby we collapsed labels across chunks associated to the same segment. We therefore repeated our analyses but instead of deriving our proportions for the Zooniverse data from the segment-level composite, we did it based on the individual chunk-level annotations. To facilitate comparison, correlations across laboratory- and Zooniverse-derived proportions are provided in Table 2.

@@

```{r tab-corage, results="as.is"}
